---
title: "Homework 5 solutions"
author: "Rio Yan"
date: '`r format(Sys.time(), "%Y-%m-%d")`'
output: github_document
---

```{r setup, include=FALSE}
library(tidyverse)

knitr::opts_chunk$set(
  fig.width = 9,
  fig.asp = .6,
  out.width = "90%"
)

theme_set(theme_minimal() + theme(legend.position = "bottom"))

options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)

scale_colour_discrete = scale_color_viridis_d
scale_fill_discrete = scale_fill_viridis_d
```


## Problem 1

Describe the raw data:we have 51 cities


Read in the data.

```{r}
homicide_df = 
  read_csv("homicide_data/homicide-data.csv") %>% 
  mutate(
    city_state = str_c(city, state, sep = "_"),
    resolved = case_when(
      disposition == "Closed without arrest" ~ "unsolved",
      disposition == "Open/No arrest" ~ "unsolved",
      disposition == "Closed by arrest" ~ "solved"
    )
  ) %>% 
  select(city_state, resolved) %>% 
  filter(city_state != "Tulsa_AL")
```


Let's look at this a bit

```{r}
aggregate_df = 
  homicide_df %>% 
  group_by(city_state) %>% 
  summarize(
    hom_total = n(),
    hom_unsolved = sum(resolved == "unsolved")
  )
```

Can I do a prop test for a single city?

```{r}
prop.test(
  aggregate_df %>% filter(city_state == "Baltimore_MD") %>% pull(hom_unsolved),
  aggregate_df %>% filter(city_state == "Baltimore_MD") %>% pull(hom_total)
) %>% 
  broom::tidy()
```

Try to iterate ...

```{r}
results_df = 
  aggregate_df %>% 
  mutate(
    prop_tests = map2(.x = hom_unsolved, .y = hom_total, ~prop.test(x = .x, n = .y)),
    tidy_tests = map(.x = prop_tests, ~broom::tidy(.x))
  ) %>% 
  select(-prop_tests) %>% 
  unnest(tidy_tests) %>% 
  select(city_state, estimate, conf.low, conf.high)
```


```{r}
results_df %>% 
  mutate(city_state = fct_reorder(city_state, estimate)) %>% 
  ggplot(aes(x = city_state, y = estimate)) +
  geom_point() +
  geom_errorbar(aes(ymin = conf.low, ymax = conf.high)) +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1))
```

## Problem 2

import one dataset 

```{r}
data_1 = read_csv("lda_data/con_01.csv")
```

```{r}
path_df = 
  tibble(
    path = list.files("lda_data"),
  ) %>% 
  mutate(
    path = str_c("lda_data/", path),
    data = map(.x = path, ~read_csv(.x)),
    path = str_remove(path, ".csv")) %>% 
  separate(path, into = c("cut", "path"), sep = "/") %>%  
  separate(path, into = c("control_arm", "subject_id"), sep = c("_")) %>% 
  select(-cut) %>% 
  unnest(data)

```

plot
```{r}
path_df_graph =
  path_df %>% 
  pivot_longer(
    week_1:week_8,
    names_to = "week_num",
    values_to = "value"
  ) %>% 
  mutate(
    control_arm = recode_factor(control_arm,"con" = "control", "exp" = "experiment")
  )

path_df_graph %>% 
  ggplot(aes(x = week_num, y = value, 
             group = subject_id, color = factor(control_arm))) +
  geom_line() + geom_point() +
  facet_grid(. ~ control_arm) +
  labs(title = "Observations on each subject between two groups",
       x = "Week Number",
       y = "Value") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

Comment on the differences between groups:
overall the graph shows that experimental group has higher value than control group, both from the start and in the end. 


## Problem 3

```{r}
set.seed(1)

sim_mean_sd = function(n, mu, sigma = 5) {
  
  sim_data = tibble(
    x = rnorm(n = 30, mean = mu, sd = sigma)
  ) 
  
  sim_data %>% 
    t.test() %>% 
    broom::tidy() %>% 
    select(estimate, p.value)
  
}

sim_results = 
  tibble(mu_list = c(0, 1, 2, 3, 4, 5, 6)) %>% 
  mutate(
    output_lists = map(.x = mu_list, ~rerun(5000, sim_mean_sd(n = 30, mu = .x))), 
    estimate_dfs = map(output_lists, bind_rows)) %>%
    unnest(estimate_dfs) %>% 
  select(-output_lists)
```

Make a plot showing the proportion of times the null was rejected (the power of the test) on the y axis and the true value of μ on the x axis. 
```{r}
reject_table = 
  sim_results %>% 
  mutate(
    decision = 
      case_when(
        p.value >= 0.05 ~ "fail to reject",
        p.value < 0.05 ~ "reject"
      )) 

reject_prop = 
  reject_table %>% 
  group_by(decision, mu_list) %>% 
  summarize(
    count = n()
  ) %>% 
  filter(decision == "reject") %>% 
  mutate(
    prop_reject = count / 5000)

prop_data1 = 
  left_join(sim_results, reject_prop, by = "mu_list")
  

prop_data1 %>% 
  ggplot(aes(x = mu_list, y = prop_reject, color = decision)) +
  geom_point() + geom_line() +
  labs(title = "Association between effect size and power",
       x = " true value of μ",
       y = "power of the test") +
  guides(color = guide_legend(title = "Prop of rejection")) 
```
Describe the association between effect size and power:



Make a second plot (or overlay on the first) the average estimate of μ̂  only in samples for which the null was rejected on the y axis and the true value of μ on the x axis. 


```{r}
mean_table = 
  sim_results %>% 
  group_by(mu_list) %>% 
  summarize(mean_mu = mean(estimate))

#plot showing average estimate of mu on the y axis and the true value of mu on x axis
plot_1 = 
  mean_table %>% 
  ggplot(aes(x = mu_list, y = mean_mu)) + 
  geom_point() + geom_line()

reject_mean = 
  reject_table %>% 
  filter(decision == "reject") %>% 
  group_by(mu_list) %>% 
  summarize(
    mean_estimate = mean(estimate)
  )

#plot second plot the average estimate in samples for which the null was rejected on y and the true value mu on x 
plot_2 =
  reject_mean %>% 
  ggplot(aes(x = mu_list, y = mean_estimate)) +
  geom_point() + geom_line()

  
#Make a plot showing the average estimate of μ̂  on the y axis and the true value of μ on the x axis. Make a second plot (or overlay on the first) the average estimate of μ̂  only in samples for which the null was rejected on the y axis and the true value of μ on the x axis. Is the sample average of μ̂  across tests for which the null is rejected approximately equal to the true value of μ? Why or why not?

combine_df = 
  left_join(mean_table, reject_mean, by = "mu_list") %>%  
  pivot_longer(
    mean_mu:mean_estimate,
    names_to = "type",
    values_to = "average_estimated_mu"
  ) %>% 
  ggplot(aes(x = mu_list, y = average_estimated_mu, color = type))+
  geom_point() + geom_line()
  

#overlay plot_1 and plot_2
  
plot(mean_table, type ="l", ylab = "mean mu",
     main = "Comparision between True mu and Average Estimated mu", xlab = "True value of mu",
     col = "blue")
par(new = TRUE)
plot(reject_mean, type = "l", xaxt = "n", yaxt = "n",
     ylab = "", xlab = "", col = "red", lty = 2)
axis(side = 4)
legend("topleft", c("true", "significant only"),
       col = c("blue", "red"), lty = c(1, 2))
```


Is the sample average of μ̂  across tests for which the null is rejected approximately equal to the true value of μ? Why or why not?














